# Examples of corpora {#corpus_example}

```{r}
library(quanteda)
library(dplyr)
library(ggplot2)
library(tidytext)
library(knitr)
```




## The Geomedia corpus

The Geomedia corpus is the result of the research project ANR Corpus Geomedia (2013-2016), coordinated by Claude Grasland (Université Paris Diderot) and Marta Severo (Univerté de Nanterre) 

The project created a tool for the harvest of RSS flows that realized collection from daily newspapers located in all countries of the world. Contrary to Mediacloud, the Geomedia project did not collected together all the feed of a newspaper but proceed to a selection of specific topics, generally limited to the *international/world* flow and the *general/breaking news* flow. But the other flows related to sport, economy, culture, ... was generaly ignored. The majority of studies realized during the project focused on the *international* feeds because they present the advantage to produce theoretically a selection of foreign news of interest for the host country where the media is located (ref.).

### The socant corpus

The "socant" corpus is a specific extraction from the Geomedia database realized in order to prepare a publication about the Migrant crisi of 2015 in the french journal *Socio-Anthropologie*. We present here a version of the corpus after several steps of cleaning described in the publication :


```{r}
qd<-readRDS("corpora/geomedia/socant.Rdata")
summary(qd,2)
head(qd,2)
```
The corpus is based on the international RSS feed of 26 newspapers published in 3 languages (French, English, Spanish) and located in 16  countries. 

```{r}
tt<-tidy(qd)

lang<-tt %>% 
      group_by(lang) %>% 
      summarise(news=n()) %>% 
      mutate(pct = 100*news/sum(news))
kable(lang, digit=1)

country<-tt %>% 
      mutate(country=substr(source,1,3))%>%
      group_by(country) %>% 
      summarise(news=n()) %>% 
      mutate(pct = 100*news/sum(news))
kable(country, digit=1)

```

Each of the source has produce different number of news through time but appears relatively regular during the period, with the exception of a short period of break of collection in January.

```{r, socant_source_time}
source_week<-tt %>%
              mutate(week=cut(as.Date(date), "weeks"))%>%
              group_by(source, week)%>%
              summarise(news=n()) %>%
              mutate(week = as.Date(week))%>%
              filter(week > as.Date("2015-01-01"))%>%
              filter(week < as.Date("2016-01-01"))

p<-ggplot(source_week, aes(x=week,y=news,color=source))+geom_line()+scale_y_log10()
p



```





## The Gazette corpus

## The Venice corpus

## The Covid corpus

